---
title: "Practical Machine Learning Project"
author: "Emerson Lima"
date: "Saturday, June 17, 2017"
---

For this dataset, the variable classe contains 5 levels:

- Class A: exactly according to the specification
- Class B: throwing the elbows to the front
- Class C: lifting the dumbbell only halfway
- Class D: lowering the dumbbell only halfway
- Class E: throwing the hips to the front

Two models will be tested using decision tree and random forest algorithms.
Cross-validation will be performed by subsampling our training dataset randomly without replacement into 2 subsamples: subTraining data (75% of the original Training data set) and subTesting data (25%).

```{r}
#Creating predictive models
#Random forest for classification and regression
#Regressive Partitioning and Regression trees
#Decision Tree plot
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
set.seed(3654)
#Loading the training data set replacing missing ("#DIV/0!") with "NA"
pmltraining <- read.csv("c:/Practical_Machine_Learning/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
#Loading the testing data set replacing missing ("#DIV/0!") with "NA"
pmltesting <- read.csv('c:/Practical_Machine_Learning/pml-testing.csv', na.strings=c("NA","#DIV/0!", ""))

dim(pmltraining)
dim(pmltesting)

#head(pmltraining)
#head(pmltesting)

#Delete columns with all missing values
pmltraining<-pmltraining[,colSums(is.na(pmltraining)) == 0]
pmltesting <-pmltesting[,colSums(is.na(pmltesting)) == 0]

#View(pmltraining)
#View(pmltesting)

# Some variables are irrelevant to our current project: 
# user_name
# raw_timestamp_part_1
# raw_timestamp_part_2 
# cvtd_timestamp 
# new_window 
# num_window 


#Columns 1 to 7 
pmltraining   <-pmltraining[,-c(1:7)]
pmltesting <-pmltesting[,-c(1:7)]


#Partitioning the training data set to allow cross-validation
#The training data set contains 53 variables and 19622 obs.
#The testing data set contains 53 variables and 20 obs.

# Cross-validation: the training data set is partionned 2 sets (subTraining (75%) and subTest (25%))

subsamples <- createDataPartition(y=pmltraining$classe, p=0.75, list=FALSE)
subTraining <- pmltraining[subsamples, ] 
subTesting <- pmltraining[-subsamples, ]

dim(subTraining)
dim(subTesting)

#View(subTraining)
#View(subTesting)


```
The variable classe contains 5 levels: A, B, C, D, E.

```{r, echo=FALSE}
plot(subTraining$classe, col="Grey", main="Levels of the variable classe (Sub-Training data set)", xlab="classe levels", ylab="Frequency")
```



```{r}
#From the graph above, we can see that each level frequency is within the same order of magnitude of each other. 
#Level A is the most frequent while level D is the least frequent

#Using Decision Tree

decision_Tree <- rpart(classe ~ ., data=subTraining, method="class")

#Predicting
prediction_DT <- predict(decision_Tree, subTesting, type = "class")

```

Plot of the Decision Tree

```{r, echo=FALSE}
rpart.plot(decision_Tree, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```

```{r}
#Test results on our subTesting data set
confusionMatrix(prediction_DT, subTesting$classe)
#Using Random Forest
random_forest <- randomForest(classe ~. , data=subTraining, method="class")

#Predicting
prediction_RF <- predict(random_forest, subTesting, type = "class")

#Test results on subTesting dataset
confusionMatrix(prediction_RF, subTesting$classe)

#Conclusion
#
#Random Forest algorithm performed better than Decision Trees.
#Accuracy for Random Forest model was 0.9941 compared to 0.7465 for Decision Tree model. 
#The expected out-of-sample error is calculated as 1 - accuracy for predictions made against the cross-validation set. 
#Our Test data set comprises 20 cases. 
#With an accuracy above 99% on our cross-validation data, we can expect that very few of the test samples will be missclassified.


#Predict Outcome levels on the original Testing data set using Random Forest

predict_final <- predict(random_forest, pmltesting, type="class")
predict_final
#Files for submission
pml_files = function(x)
{
  n = length(x)
  for(i in 1:n)
  {
    filename = paste0("problem_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_files(predict_final)
```

